The Way Forward in Parallelizing Dynamic Languages

### 2014-05-01 talk/icooolps2014/position-paper.tex
shared-nothing draft

- often needs major restructuring of programs (explicit data exchange)\\
- sometimes communication overhead is too large\\
- shared memory is a problem, copies of memory are too expensive

>>>>

There are also approaches that work around the GIL instead of trying
to replace it. If an application can be split into completely
independent parts that only very rarely need to share anything, or
only do so via an external program like a database, then it is
sensible to have one GIL per independent part. As an example of such
an approach we look at the
\emph{multiprocessing}\footnote{https://docs.python.org/2/library/multiprocessing.html}
module of Python. In essence, it uses process-forking to provide the
application with multiple interpreters that can run in parallel.
Communication is then done explicitly through pipes.

Obviously not every application fits well into this model and its
applicability is therefore quite limited. Performance is good as
long as the application does not need to communicate a lot, because
inter-process communication is relatively expensive. Also the
implementation of this approach is very cheap since one can
actually take an unmodfied GIL-supported interpreter and run
multiple of them in parallel. That way, we also inherit the
easy integration of external libraries without any changes.
While the model of explicit communication is often seen as a
superior way to synchronize concurrent applications because
of its explicitness, it does not actually introduce a better
synchronization mechanism for applications.

%% - often needs major restructuring of programs (explicit data exchange)\\
%% - sometimes communication overhead is too large\\
%% - shared memory is a problem, copies of memory are too expensive
